### 2025-03-14 literature.qmd
initialise

1) Knight, J., Shi, L., & Haiyuan, W. (2022). Why has China’s Inequality of Household Wealth Risen Rapidly in the Twenty‐First Century?. Review of Income and Wealth, 68(1), 109-138.

This paper tries to explain the rise in wealth inequality in 21st century China. It used the CHIP data in 2002 and 2013, emphasizing that these two releases are rigorously comparable. It attempts to explain the rise in terms of differential saving rate, differential house price inflation, income from wealth and a growing urban-rural wealth disparity. It found that the high saving rate of the rich, the housing boom, especially for the already expensive houses, real capital gains from houses, and urban-rural wealth disparity are the main causes for the rising wealth inequality, while the impact of income from wealth excluding real capital gain is marginal. This is a very important citation since it contains many technical details for wealth inequality studies. * Since housing value is missing in CHIP and housing is over 50% of net wealth for Chinese households, I abandoned using it and pivoted to the CHFS data (China Household Finance Program by a big university) instead.

2) Zhang, M. L., Brooks, M. R., Ding, D., Ding, H., He, H., Lu, J., & Mano, R. (2018). China’s high savings: drivers, prospects, and policies. International Monetary Fund.

This paper provides a comprehensive empirical review of China’s saving rates across sectors. It offers some account of historical events and their impact on saving rates. It briefly talks about the relation between income inequality and savings inequality. It mentions some fancy terms that I could reference in my paper, such as financial repression, which could be related to income and wealth inequality. It also has some regression examples with Chinese characteristics that I can learn from, such as using provincial dummies and SOE dummies. In sum, this paper provides a good example of how to construct historical narratives for empirical analysis.

3) Piketty, T., Yang, L., & Zucman, G. (2019). Capital accumulation, private property, and rising inequality in China, 1978–2015. American Economic Review, 109(7), 2469-2496.

This paper utilised various data sources to estimate the state of inequality in China. It constructed the asset side of the national balance sheet of China from 1978 to 2015, which shows the growth and composition of national wealth. It has the Piketty-style description of inequality comparing the income and wealth equality across groups. It has data for the ownership structure of China’s national wealth and corporations. There are rich comparisons between China and other countries. I could cite the data directly from this paper or reproduce certain graphs using the CHFS data. Since this paper uses various sources of data, I could cite the results in this paper to confirm or confront the results in my paper. This paper is an excellent source of background information for the literature review section of my paper.

4) Kanbur, R., Wang, Y., & Zhang, X. (2021). The great Chinese inequality turnaround. Journal of Comparative Economics, 49(2), 467-482.

This paper counters the popular view that China’s inequality has still been rising in recent years and argues that China’s inequality has plateaued. This paper used CHIP and CFPS data. It has detailed documentation and decomposition of Gini index and generalised entropy indexes, and it shows that China’s inequality peaked around 2010. It has an interesting decomposition of Theil index into within-group and between-group components. The decomposition shows that much of the Theil’s T is contributed by the urban-rural divide, which significantly shrank after 2010. The coastal-inland divide was small before the reforms, but it also contributed almost 0.05 to the overall Theil’s T in the mid 2000s, which has been declining since then. The results are counter-intuitive to me as the within urban-rural contribution is tiny compared to the between-urban-rural contribution. Another counter-intuitive observation is that the authors use two different sources and get vastly different Gini and Theil indexes. Regarding my paper, I can decompose within-province and between-province. The authors attributed the moderation of inequality to government policies. The within-between group decomposition seems to be novel and easy to implement. This paper talks about the Lewis turning point and the Kuznets turning point, which I am not familiar with.

5) Schettino, F., Gabriele, A., & Khan, H. A. (2021). Polarization and the middle class in China: A non-parametric evaluation using CHNS and CHIP data. Structural Change and Economic Dynamics, 57, 251-264.

This paper studies the polarization of China and its impact on the middle class. Polarization is an interesting intersectional idea of economics and sociology. The idea suggests that with rising polarization, the alienation between groups in a society worsens, which might cause social conflicts. This paper used Relative Distribution to distinguish between pure growth effects and pure distributional effects, and examine how the two effects have influenced the structure and size of the national middle class. This paper found that between 2000 and 2015, rapid economic growth of China lifted most segments of the population (location effect), and led to a rise in households joining the “global middle class”, but the “shape effects” or the “hollowing out” of the middle-income groups occurred. Interesting way of decomposing distribution changes into pure growth and pure distribution. I can cite their results or replicate their results with the new 2018 data. I have no idea why they did not use the 2018 data, given they wrote this in 2021.

6) Zhang, J. (2021). A survey on income inequality in China. Journal of Economic Literature, 59(4), 1191-1239.

This is a comprehensive review of works on China’s inequality. It’s a good resource to look for citations. Its findings resemble 4) that inequality in China has been rising until the 2010s when the trend slightly reversed. The trends of urban-rural disparity and regional inequality are closely correlated with the overall inequality from the 1980s to the 2000s, but no longer after the 2010s. It reported a negative correlation between inequality and intergenerational mobility, consistent with the Great Gatsby Curve in developed countries.

7) Mookherjee, D., & Shorrocks, A. (1982). A decomposition analysis of the trend in UK income inequality. The Economic Journal, 92(368), 886-902.

This is a classical piece on how to decompose inequality into within-group and between-group effects. It also has examples on how to present the decomposed effects in a chronological way. My paper might reference this paper for methodologies.

8) Liu, Y., Hao, Y., & Lu, Z. N. (2022). Health shock, medical insurance and financial asset allocation: evidence from CHFS in China. Health Economics Review, 12(1), 52.

Since I have decided no longer to use the CHIP data, I need to find some examples of using CHFS data. This is a study of how various types of health insurance help households deal with health shocks using the CHFS data. It also talks about the relation between health insurance and financial mobility. It has some examples of dividing assets into safe and risky assets. I could reference this paper for technical details.

9) Zeng, T., & Zhu, S. (2022). The mobility of top earnings, income, and wealth in China: Facts from the 2011–2017 China household finance survey. Journal of Asian Economics, 80, 101461.

This is a paper that uses the CHFS data to study wealth inequality and inter-generational mobility. It has many examples of presenting income and wealth compositions of different income groups. It noticed that the top 1% allocated a much higher proportion of wealth to private equity, which also fluctuates a lot. I might check if my results are similar to theirs if I also start from the year 2011.

10) Tan, J., Zeng, T., Zhu, S., D’ambrosio, C., Gan, L., He, H., & Zhang, J. (2017). Earnings, income, and wealth distributions in China: Facts from the 2011 China Household Finance Survey. Working Paper.

Another example of using the CHFS data to study inequality. This paper found that the level of inequality in 2011 was comparable to that of the U.S. in 1928. It has many graphs and technical details. I can use it in the literature review and in the data part.

11) Gan, L., Yin, Z., & Tan, J. (2016). Report on the development of household finance in rural China (2014). Springer Singapore.

This is a paper written by the project team of CHFS. It has a very detailed introduction of the data, including information such as the refusal rate, household composition, and gender structure. It then went on with a highly detailed report of the development of household finance in rural China. I will cite this paper for technical details of the data and some statistics.

12) Xie, Y., & Jin, Y. (2015). Household wealth in China. Chinese sociological review, 47(3), 203-229.

Another more cited empirical paper on China’s wealth inequality but it did not use CHFS. I can reference it for facts in literature review. These last three papers are basically presenting data and have less to do with inequality measure decomposition. The CHFS dataset is huge with thousands of variables, and thus it is important to follow how they aggregate individual asset / liability items into broad asset classes.