### 2025-04-23 report/draft.qmd
update

---
title: "Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS"
format: 
    pdf:
        documentclass: article 
        number-sections: true
        margin-top: 1in
        margin-bottom: 1in
        margin-left: 1in
        margin-right: 1in
        linestretch:  1.2
        fontsize: 10pt
        fig-pos: "H"
        author: "Chenning Xu"
bibliography: tom.bib
---

# Introduction
China has experienced remarkable economic growth in recent decades following its reform and opening-up policy initiated in 1978. However, this period has also been marked by a significant rise in income and wealth inequality [@piketty_capital_2019]. Some scholars have provided evidence suggesting that China's overall inequality peaked around 2010 and subsequently began to decline [@kanbur_great_2021; @zhang_survey_2021]. It is also well-documented that China's inequality is strongly intertwined with its urban-rural and coastal-inland divisions [@piketty_capital_2019; @zhang_survey_2021]. This paper utilizes data from the 2017 and 2019 China Household Finance Survey (CHFS) to examine the current state of China's wealth inequality. Specifically, it investigates whether the latest publicly available data supports the view that China's inequality has continued to decline since around 2010, as well as how China's wealth inequality relates to urban-rural and regional disparities.

# Background
This section presents some notable findings on China's income and wealth inequality in the existing literature.

China has experienced a considerable increase in income and wealth inequality since the reform and opening-up policy initiated in 1978. Using diverse sources, including tax records, household surveys, and private wealth rankings, @piketty_capital_2019 provide a detailed analysis of the transformation in China's inequality profile. Their findings show that between 1978 and 2015, the income share of the top 10% increased from 27% to 41%, while the share of the bottom 50% decreased from 27% to 15%. Wealth concentration also intensified significantly: the wealth share held by the top 10% rose from around 40% to nearly 70%, whereas the bottom 50%'s share dropped from over 15% to about 5%. Additionally, they documented an increase in the urban-to-rural per capita income ratio, from less than 200% in 1978 to approximately 350% by 2015. Humorously noting this stark shift, they remarked that China had transitioned from Nordic-level equality to U.S.-level inequality during the reform era.

@knight_why_2022 further document a rapid rise in China's household wealth inequality between 2002 and 2013, reporting that the national Gini coefficient for household wealth per capita increased markedly from 0.50 to 0.62. They identify housing wealth as the primary driver behind this increase, with its share rising from 53% to 73% of total household wealth, and its contribution to overall wealth inequality growing from 64% to 79%. The authors emphasize that differential savings rates across income groups, uneven house price inflation—particularly pronounced in major urban centers—and widening urban-rural disparities significantly intensified wealth inequality during this period.

The income and wealth inequality highly intersects with urban-rural and reginal disparities. @zhang_survey_2021 highlights significant rural-urban and regional disparities underlying China's rising inequality since its economic reforms. The urban-rural income ratio increased sharply, peaking at around 3.3 in 2009 before declining slightly to 2.7 by 2019, still substantially higher than pre-reform levels. Despite a notable decline in overall inequality after 2010—with the national Gini coefficient decreasing from a peak of 0.491 in 2008 to approximately 0.465 in 2019—regional inequality remains pronounced. For instance, per capita GDP in the eastern coastal region was roughly 2.6 times higher than in the western region by 2018. Additionally, rural-to-urban migration grew dramatically from fewer than 20 million in 1990 to approximately 290 million by 2019, reflecting both the economic disparity between regions and the persistent urban-rural divide, exacerbated by restricted access to urban public services under the Hukou household registration system.

Consistent with the finding in @zhang_survey_2021 that China's inequality peaked in 2008 and 2009, @kanbur_great_2021 similarly document that China's inequality increased sharply after economic reforms but started to plateau around 2010. They find the national Gini coefficient peaked at about 0.525 in 2010, subsequently declining to approximately 0.476 in 2016 before rising slightly again. Consistent with Zhang (2021), they attribute this recent turnaround primarily to narrowing rural-urban and coastal-inland disparities, driven by demographic changes, tightening rural labor markets, and rising rural wages.

# Data
The analysis in this study utilizes data from the 2017 and 2019 waves of the China Household Finance Survey (CHFS), a nationally representative dataset collected by the Survey and Research Center for China Household Finance at Southwestern University of Finance and Economics [@gan_report_2014]. The CHFS dataset includes household-level financial information on income, assets, debts, consumption, and demographic characteristics. Data are collected through face-to-face interviews using a multi-stage probability proportional to size (PPS) sampling method, ensuring representativeness at both national and provincial levels. Sample weights provided by CHFS are calculated to account for the sampling design, adjusting for unequal probabilities of selection at different stages.

Several limitations of the CHFS dataset should be noted. First, it relies on self-reported values for income and assets, potentially introducing biases such as underreporting, recall inaccuracies, or privacy concerns [@gan_report_2014]. Second, definitional changes across survey waves may affect comparability. For instance, the 2019 survey separately records garage assets, previously included within housing assets in the 2017 wave, potentially complicating longitudinal analyses. Additionally, certain variables in the dataset are top-coded to protect respondent privacy, with extreme values replaced by preset caps, accompanied by censoring indicators [@chfs_2019_2019]. Such censoring is not accounted for in this study and might underestimate the level of income and wealth inequality in China.

The CHFS dataset has been extensively utilized in academic research. Studies leveraging the CHFS data have explored topics ranging from wealth distribution and inequality [@tan_earnings_2017], household finance in rural China [@gan_household_2014], household medical insurance expenditure [@li_household_2019], and many others.

::: {#tbl-tbl1}

{{< include ../output/summary_stats.txt >}}

All figures are reported in RMB at constant 2017 prices.

Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.

Mean, Median, and Gini Coefficient for Income and Wealth in China, 2011, 2017, and 2019
:::

Compared to the 2011 data, the Gini coefficient for income has decresed from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019. The Gini coefficient for wealth has decreased from 0.761 in 2011 to 0.701 in 2017 and continued to decrease to 0.667 in 2019. The results show that China's income and wealth inequality has been decreasing since 2011. The finding is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021, that China's inequality peaked around 2010 and started to decrease since then.

Notably, the growth rate of mean and median income from 2011 to 2017 is clearly higher than that of wealth but from 2017 to 2019, the growth of mean and median wealth surpasses that of income by a great margin.

## Factor decomposition of wealth inequality

To understand the sources driving wealth inequality in China, this section applies the decomposition method developed by @lerman_income_1985. According to this method, the contribution of each wealth component to the overall Gini coefficient is decomposed as follows:

$$
G = \sum_{k=1}^{K} S_k G_k R_k
$$

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking [@lerman_income_1985].

::: {#tbl-wealth-summary}

{{< include ../output/wealth_summary.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary Statistics of Wealth, 2017 and 2019
:::

@tbl-wealth-summary summarizes household wealth allocation in China for 2017 and 2019. Housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types. Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets. Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.

Financial equity notably increased from 2017 to 2019, with mean values rising substantially from 103,355 RMB to 161,199 RMB and median values more than doubling from 17,170 RMB to 39,251 RMB. Correspondingly, financial assets increased their share in total household wealth from around 12% in 2017 to approximately 15% in 2019. Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate]. Observed changes in land, commercial, and other assets should be interpreted cautiously, as it appears that some assets are redefined or newly added to the variables, which might partly reflect definitional adjustments rather than purely economic dynamics.

::: {#tbl-gini-decomposition-2017}

{{< include ../output/2017_gini_decom.txt >}}

Source: Author's calculation based on the 2017 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2017
:::

::: {#tbl-gini-decomposition-2019}

{{< include ../output/2019_gini_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019
:::

@tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively. Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality. This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011. Financial equity emerged as another important factor, increasing its contribution from approximately 11% in 2017 to about 14% in 2019. In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.

Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019. Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors. The following part expands this analysis by conducting group-specific wealth inequality decompositions to understand the difference in wealth inequality across different rural and urban households and east and non-east regions.

## Factor decomposition partitioned by rural and urban residence
Since changes in the factor decomposition from 2017 to 2019 for rural and urban households closely mirror those observed for the entire population, only the 2019 decomposition results for rural and urban groups are presented here.

::: {#tbl-urban-decom}

{{< include ../output/2019_urban_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Urban)
:::

::: {#tbl-rural-decom}

{{< include ../output/2019_rural_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Rural)
:::

The comparative factor decompositions for 2019 underscore several salient distinctions between the determinants of wealth inequality in urban and rural China.  In both settings residential property remains the predominant source of inequality; nevertheless, its influence is markedly stronger in cities, where housing contributes roughly three‑quarters of the overall Gini coefficient, compared with just over one‑half in the countryside.  This pronounced urban effect reflects the exceptionally high and spatially differentiated prices of metropolitan real estate as well as the tendency for affluent urban households to own multiple dwellings.

By contrast, assets that are either absent from, or less significant in, urban portfolios assume a more prominent role in rural areas.  Most notably, agricultural and homestead land accounts for almost one‑fifth of rural inequality but for only about six per cent of the urban total.  The unequal distribution of land holdings, combined with large inter‑village variation in implicit land values, therefore constitutes a key driver of rural wealth dispersion.  Commercial (business) assets exhibit a similar pattern: they explain roughly eleven per cent of rural inequality but just seven per cent in cities, mirroring the heterogeneity of small‑scale entrepreneurial activities in village and township economies.  Conversely, financial equity is slightly more important in urban areas (around thirteen per cent) than in rural areas (about eleven per cent), consistent with better access of urban households to formal financial markets.

Finally, debt liabilities play a modest equalising role in both residence types, as indicated by their negative contributions to the Gini.  The mitigating effect is somewhat larger in urban China: mortgage debt alone reduces the urban coefficient by roughly 3.3 percentage points, relative to 2.1 percentage points in rural areas.  Taken together, these findings imply that policy initiatives aimed at curbing wealth inequality ought to be tailored to local contexts.  In cities, measures that temper speculative housing demand and widen access to diversified financial products are likely to prove most effective, whereas in rural regions reforms that strengthen land‑tenure security, facilitate land transactions, and support productive agricultural and non‑farm enterprises are likely to yield greater distributive gains.

When benchmarked against the national decompositions for 2017 and 2019 (see \@tbl-gini-decomposition-2017 and \@tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.  What the residence‑specific results add, however, is a clearer sense of _relative_ importance: housing is even more dominant in cities than the country‑wide average, whereas land and business assets loom larger in rural portfolios than the aggregate figures would suggest.  At the same time, the contribution of financial assets in both residence types remains close to the national share, confirming that the rise of household financial investment is a broad‑based rather than location‑specific phenomenon.

## Factor decomposition partitioned by east and non-east regions
It is worth noting that from 2017 to 2019, housing's contribution to the Gini coefficient rose by 2.4 percentage points among households in the East region, whereas it decreased by 2.9 percentage points among Non-East households, despite housing's overall contribution remaining nearly identical across both years. This finding is significant, indicating that housing-related wealth inequality increased in East China but decreased in Non-East regions during this period. Aside from housing, contributions from other asset and debt categories showed similar trends for East and Non-East households. Therefore, only the decomposition results for 2019 are presented here.

::: {#tbl-east-decom}

{{< include ../output/2019_east_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (East)
:::

::: {#tbl-noneast-decom}

{{< include ../output/2019_noneast_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Non-East)
:::

Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China.  Yet they also reveal distinct regional configurations.  In the coastal East, the contribution of housing climbs to more than 80 per cent of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets.  By contrast, in the Non‑East provinces housing accounts for just 58 per cent, while land and commercial (business) assets together explain nearly one‑quarter of inequality, double their combined share in the East.  Financial equity is a sizeable driver in both regions, but its influence is marginally greater in the interior (≈ 16 %) than along the coast (≈ 12 %), reflecting the recent expansion of retail investment beyond the major coastal cities.  Debt components continue to dampen inequality everywhere; however, the equalising effect of mortgage liabilities is slightly stronger in the East, where household borrowing is more closely tied to high‑value housing.  These contrasts suggest that policies aimed at moderating wealth disparities will need to account for regional heterogeneity—addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within the common framework of wider financial inclusion.

## Group decomposition of wealth inequality by rural and urban
To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967]. The overall Gini coefficient  is decomposed as follows:

$$G = G_{B} + \sum_{k} a_{k}G_{k} + R$$

where $G_{B}$ is the between-group Gini, which captures inequality arising solely from differences in group means; $G_{k}$ is the within-group component, a weighted sum of the inequality within each group, with weights $a_{k}$ representing the products of each group's population share and wealth share; and $R$ is the overlap (residual) component, which captures inequality due to the intersection of group wealth distributions. Lambert and Aronson (1993) provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap. A positive $R$ thus indicates significant intersections among wealth distributions of different groups[@lambert_inequality_1993].

::: {#tbl-tbl4}

{{< include ../output/rural_urban_summary.txt >}}

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary statistics of urban and rural income and wealth, 2017 & 2019
:::

The summary statistics highlight significant and persistent disparities between urban and rural China. Urban households consistently exhibit higher mean and median incomes and wealth compared to rural households. Urban mean wealth was 3.48 times greater than rural mean wealth in 2017 and widened slightly to 3.78 times in 2019. Median urban wealth was 2.82 times higher than the rural median in 2017, though this ratio decreased slightly to 2.52 in 2019. Additionally, the Gini coefficient for wealth is consistently higher in rural areas, though both rural and urban Gini coefficients declined modestly between 2017 and 2019.

::: {#tbl-urbanrural-decomp}

{{< include ../output/urban_rural_gini_decomp.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Decomposition of wealth Gini coefficient by urban–rural division, 2017 & 2019 (%)
:::

The decomposition results indicate that about 55–57% of total wealth inequality is attributable to within-group disparities among urban and rural households. The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34%. The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% of total inequality. The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.

::: {#tbl-region-summary}

{{< include ../output/region_summary.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary Statistics by Region, 2017 and 2019
:::

The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019. Notably, the mean and median wealth of the West-area households surpassed that of the Middle in 2019, but West's Gini coefficient rose to the highest among all regions in 2019 as well.

::: {#tbl-region-decomp}

{{< include ../output/region_gini_decomp.txt >}}

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Decomposition of Wealth Gini Coefficient by Region, 2017 and 2019 (%)
:::

The group decompsition shows that similar to the urban-rural decomposition, the between-group component contribues over 30% to the overall wealth inequality in both 2017 and 2019. The between group contribution saw a slight decrease from 2017 to 2019 with a similari increase in the residual component. This might signal some modest convergence in household wealth between regions.

# References