### 2025-04-25 report/draft.qmd
update

---
title: "Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS"
format: 
    pdf:
        documentclass: article 
        number-sections: true
        margin-top: 1in
        margin-bottom: 1in
        margin-left: 1in
        margin-right: 1in
        linestretch:  1.2
        fontsize: 10pt
        fig-pos: "H"
        author: "Chenning Xu"
bibliography: tom.bib
---

# Introduction

China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy. This period yielded high average GDP growth rates alongside considerable improvements in poverty reduction, health, and education [@world_bank_world_2024; @ravallion_chinas_2009]. This growth, however, manifested unevenly across time, geography, and economic sectors. Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016]. Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024]. China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].

Parallel to this economic expansion, China experienced a substantial increase in wealth and income inequality. Analysis combining diverse data sources reveals a marked shift in distribution between 1978 and 2015; the wealth share controlled by the top 10% increased from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished considerably [@piketty_capital_2019]. Wealth inequality escalated rapidly, with the Gini coefficient for household wealth per capita rising markedly between 2002 and 2013, reaching high levels by international comparison [@knight_why_2022; @tan_earnings_2017]. Some research suggests that overall income inequality may have reached a peak around 2008-2010, potentially followed by a gradual decline or stabilization, linked partly to narrowing structural divides [@zhang_survey_2021; @kanbur_great_2021]. It is well-documented that China's inequality dynamics intersect closely with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021]. The urban-rural income ratio remained high even in the late 2010s, despite some narrowing from its 2009 peak, and regional income differences persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].

This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS). The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income. The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality. A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions. Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups. This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.

# Literature Review
This section reviews notable findings from existing literature on the evolution of China's income and wealth inequality following the market-oriented reforms initiated in 1978.

China's economic trajectory since 1978 has been characterized by high rates of growth alongside a considerable increase in economic disparities. Utilizing diverse data including tax records, surveys, and wealth rankings, @piketty_capital_2019 meticulously document this transformation. Their analysis reveals a substantial redistribution of national income between 1978 and 2015: the top 10% saw their income share rise from 27% to 41%, while the bottom 50%'s share conversely fell from 27% to 15%. Wealth concentration followed a similar, potentially more pronounced, pattern. The wealth share of the top 10% increased from approximately 40% to nearly 70%, diminishing the bottom 50%'s share from over 15% to around 5%. Concurrently, the urban-rural income gap widened, with the average per capita income ratio increasing from less than 2:1 to about 3.5:1. This shift led @piketty_capital_2019 to observe that China moved from equality levels comparable to Nordic countries towards those observed in the U.S. during this period. Complementing this, @xie_income_2014 established, using multiple surveys from the mid-2000s onwards, that China's Gini coefficient had reached levels between 0.53 and 0.55, exceeding official figures and the US benchmark at the time.

Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62. A key finding is the role of housing wealth, which grew from 53% to 73% of total household assets and accounted for an increasing share of wealth inequality, rising from 64% to 79% of the total. The authors attribute the intensification of wealth inequality during these years to a combination of factors, including differential saving rates favoring the wealthy, and uneven house price inflation which benefited homeowners in major urban centers.

These trends in income and wealth inequality are deeply intertwined with persistent structural divides within China. Significant rural-urban and regional disparities remain central to understanding China's inequality landscape [@zhang_survey_2021; @xie_income_2014]. The urban-rural income ratio, despite some decline after peaking around 3.3 in 2009, remained elevated at approximately 2.7 in 2019 [@zhang_survey_2021]. This gap is reinforced by the *hukou* (household registration) system, which limits rural migrants' access to urban services and opportunities, even as migration increased substantially from 1990 to 2019. Regional inequality also persists, with coastal areas generally outpacing inland regions [@liao_sixty_2016]. While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@zhang_survey_2021; @kanbur_great_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.

The transition towards a market economy itself introduced multiple drivers of inequality [@benjamin_income_2008; @zhang_survey_2021]. In rural China, the rise was associated with unequal access to non-agricultural income sources, particularly family businesses, alongside periods of slow growth or falling prices in the agricultural sector [@benjamin_income_2008; @ravallion_chinas_2009]. In urban areas, contributing factors included the phasing out of subsidies and entitlements, increasing wage dispersion driven by market forces and enterprise reforms, higher returns to education (especially tertiary degrees), and unemployment resulting from State-Owned Enterprise (SOE) restructuring [@benjamin_income_2008]. Uneven regional development was further amplified by policies favoring coastal regions, as well as the differential impacts of globalization and marketization [@liao_sixty_2016]. It is worth noting, however, that considerable inequality also exists within specific localities, suggesting that factors beyond broad regional or urban-rural classifications are also influential [@benjamin_income_2008].

Furthermore, the pattern of economic growth influenced distributional outcomes. Research indicates that growth originating in the agricultural and rural sectors was more effective at reducing both poverty and overall inequality compared to growth in the industrial or service sectors [@ravallion_chinas_2009]. This suggests the importance of the initial economic structure and the specific mechanisms linking growth to income gains across population segments. Conversely, the rise in inequality itself appeared to impede poverty reduction; provinces that began the reform period with higher levels of inequality experienced slower subsequent progress in reducing poverty, partly because growth was less effective at lifting incomes at the lower end of the distribution in these areas [@ravallion_chinas_2009]. Specific policies, such as the initial agrarian reforms and adjustments to agricultural procurement prices, also demonstrably affected poverty and inequality levels. While income inequality may have plateaued or slightly decreased post-2010, linked to narrowing structural gaps and potentially policy interventions [@zhang_survey_2021; @liao_sixty_2016], wealth concentration, influenced by real estate appreciation, continued its upward trend well into the 2010s [@knight_why_2022; @piketty_capital_2019].

# Data
The empirical analysis in this paper utilizes microdata from the 2017 and 2019 waves of the China Household Finance Survey (CHFS). The CHFS, conducted biennially by the Survey and Research Center for China Household Finance at Southwestern University of Finance and Economics (SWUFE) since 2011, is designed to collect household financial information representative at both national and provincial levels [@gan_report_2014]. The survey employs a stratified, three-stage Probability Proportional to Size (PPS) random sampling methodology. This approach involves selecting counties, followed by village or neighborhood committees, and then households, incorporating oversampling of high-wealth households to better capture asset distribution.

Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels. As with most survey data, CHFS figures are subject to potential limitations. Asset and income data rely on self-reporting, which can be affected by reporting error or bias. Respondent privacy is protected through methods including the top-coding of certain high-value income and asset variables; this may result in an underestimation of the upper tails of the respective distributions and potentially impact inequality metrics. Researchers should also note that minor adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019, may require consideration for longitudinal comparisons. Missing data are handled through specific coding and imputation for certain variables.

The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis. The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior. Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others [@tan_earnings_2017; @gan_report_2014; @liu_health_2022].

The 2017 CHFS wave surveyed approximately 38,994 households, while the 2019 wave included 34,643 households across 343 county-level units in 29 provinces [@chfs_2017_2017; @chfs_2019_2019]. For the present analysis, observations identified as low quality in the 2017 data, along with those lacking essential variables for this study, were omitted; however, the final analytical sample retains the substantial majority of the original observations.

# Analysis
## Summary Statistics
::: {#tbl-tbl1}

{{< include ../output/summary_stats.txt >}}

All figures are reported in RMB at constant 2017 prices.

Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.

Mean, Median, and Gini Coefficient for Income and Wealth in China, 2011, 2017, and 2019
:::

Compared to the 2011 data, the Gini coefficient for income has decresed from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019. The Gini coefficient for wealth has decreased from 0.761 in 2011 to 0.701 in 2017 and continued to decrease to 0.667 in 2019. The results show that China's income and wealth inequality has been decreasing since 2011. The finding is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021, that China's inequality peaked around 2010 and started to decrease since then.

Notably, while the growth rate of mean and median income from 2011 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, where the growth of mean and median wealth greatly surpassed income growth. This later-period divergence may reflect factors such as significant asset price appreciation relative to income gains.

## Factor decomposition of wealth inequality

To understand the sources driving wealth inequality in China, this section applies the decomposition method developed by @lerman_income_1985. According to this method, the contribution of each wealth component to the overall Gini coefficient is decomposed as follows:

$$
G = \sum_{k=1}^{K} S_k G_k R_k
$$

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking [@lerman_income_1985].

::: {#tbl-wealth-summary}

{{< include ../output/wealth_summary.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary Statistics of Wealth, 2017 and 2019
:::

@tbl-wealth-summary summarizes household wealth allocation in China for 2017 and 2019. Housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types. Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets. Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.

Financial equity notably increased from 2017 to 2019, with mean values rising substantially from 103,355 RMB to 161,199 RMB and median values more than doubling from 17,170 RMB to 39,251 RMB. Correspondingly, financial assets increased their share in total household wealth from around 12% in 2017 to approximately 15% in 2019. Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate]. However, later results in this paper show that the gini coefficient for land in the rural area was 89.7% in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas. This shows that dispite a wide distribution of land ownership, the distribution of land is highly unequal in modern China.

Observed changes in land, commercial, and other assets should be interpreted cautiously, as it appears that some assets are redefined or newly added to the variables, which might partly reflect definitional adjustments rather than purely economic dynamics.

::: {#tbl-gini-decomposition-2017}

{{< include ../output/2017_gini_decom.txt >}}

Source: Author's calculation based on the 2017 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2017
:::

::: {#tbl-gini-decomposition-2019}

{{< include ../output/2019_gini_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019
:::

@tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively. Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality (72.7% in 2017, 72.3% in 2019). This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011. Financial equity emerged as another important factor, increasing its contribution from approximately 11.5% in 2017 to about 13.5% in 2019. In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.

Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019. Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors. To explore heterogeneity in these patterns, the following sections decompose wealth inequality within specific demographic subgroups.

## Factor decomposition partitioned by rural and urban residence
Since changes in the factor decomposition from 2017 to 2019 for rural and urban households closely mirror those observed for the entire population, only the 2019 decomposition results for rural and urban groups are presented here.

::: {#tbl-urban-decom}

{{< include ../output/2019_urban_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Urban)
:::

::: {#tbl-rural-decom}

{{< include ../output/2019_rural_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Rural)
:::

The comparative factor decompositions for 2019 underscore several salient distinctions between the determinants of wealth inequality in urban and rural China.  In both settings residential property remains the predominant source of inequality; nevertheless, its influence is markedly stronger in cities, where housing contributes roughly three‑quarters of the overall Gini coefficient, compared with just over one‑half in the countryside.  This pronounced urban effect reflects the exceptionally high and spatially differentiated prices of metropolitan real estate as well as the tendency for affluent urban households to own multiple dwellings.

By contrast, assets that are either absent from, or less significant in, urban portfolios assume a more prominent role in rural areas.  Most notably, agricultural and homestead land accounts for almost one‑fifth of rural inequality but for only about six per cent of the urban total.  The unequal distribution of land holdings, combined with large inter‑village variation in implicit land values, therefore constitutes a key driver of rural wealth dispersion.  Commercial (business) assets exhibit a similar pattern: they explain roughly eleven per cent of rural inequality but just seven per cent in cities, mirroring the heterogeneity of small‑scale entrepreneurial activities in village and township economies.  Conversely, financial equity is slightly more important in urban areas (around thirteen per cent) than in rural areas (about eleven per cent), consistent with better access of urban households to formal financial markets.

Finally, debt liabilities play a modest equalising role in both residence types, as indicated by their negative contributions to the Gini.  The mitigating effect is somewhat larger in urban China: mortgage debt alone reduces the urban coefficient by roughly 3.3 percentage points, relative to 2.1 percentage points in rural areas.  Taken together, these findings imply that policy initiatives aimed at curbing wealth inequality ought to be tailored to local contexts.  In cities, measures that temper speculative housing demand and widen access to diversified financial products are likely to prove most effective, whereas in rural regions reforms that strengthen land‑tenure security, facilitate land transactions, and support productive agricultural and non‑farm enterprises are likely to yield greater distributive gains.

When benchmarked against the national decompositions for 2017 and 2019 (see @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.  What the residence‑specific results add, however, is a clearer sense of _relative_ importance: housing is even more dominant in cities than the country‑wide average, whereas land and business assets loom larger in rural portfolios than the aggregate figures would suggest.  At the same time, the contribution of financial assets in both residence types remains close to the national share, confirming that the rise of household financial investment is a broad‑based rather than location‑specific phenomenon.

## Factor decomposition partitioned by east and non-east regions
It is worth noting that from 2017 to 2019, housing's contribution to the Gini coefficient rose by 2.4 percentage points among households in the East region, whereas it decreased by 2.9 percentage points among Non-East households, despite housing's overall contribution remaining nearly identical across both years. This finding is significant, indicating that housing-related wealth inequality increased in East China but decreased in Non-East regions during this period. Aside from housing, contributions from other asset and debt categories showed similar trends for East and Non-East households. Therefore, only the decomposition results for 2019 are presented here.

::: {#tbl-east-decom}

{{< include ../output/2019_east_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (East)
:::

::: {#tbl-noneast-decom}

{{< include ../output/2019_noneast_decom.txt >}}

Source: Author's calculation based on the 2019 CHFS data.

Factor Decomposition of Wealth Gini Coefficient, 2019 (Non-East)
:::

Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China, yet they also reveal distinct regional configurations. In the coastal East, housing's influence is particularly pronounced; its contribution climbs to over 80 per cent (80.3%) of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets and its strong alignment with overall wealth rank (Gini correlation 0.96). By contrast, in the Non‑East provinces housing accounts for a lower share at 58.3 per cent. Here, land and commercial (business) assets play a much larger role, together explaining nearly one-quarter of inequality (23.2%, combining 13.8% from land and 9.4% from commercial), substantially higher than their combined share of about 7.6% in the East. This reflects not only portfolio differences but also stronger correlations between these assets and wealth rank in the Non-East (Gini correlations of 0.83 for land and 0.80 for commercial, vs. 0.35 and 0.72 respectively in the East).

Financial equity is a sizeable driver in both regions, though its contribution and correlation with wealth rank are marginally greater in the interior (15.7% contribution, 0.78 correlation) than along the coast (12.4% contribution, 0.77 correlation), perhaps reflecting the recent expansion of retail investment beyond major coastal cities. Debt components, particularly mortgage liabilities, serve to dampen inequality in both regions through their negative contributions to the overall Gini coefficient. These contrasts underscore that policies aimed at moderating wealth disparities need to account for regional heterogeneity—potentially addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within a common framework of wider financial inclusion.

## Group decomposition of wealth inequality by rural and urban
To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967]. The overall Gini coefficient  is decomposed as follows:

$$G = G_{B} + \sum_{k} a_{k}G_{k} + R$$

where $G_{B}$ is the between-group Gini, which captures inequality arising solely from differences in group means; $G_{k}$ is the within-group component, a weighted sum of the inequality within each group, with weights $a_{k}$ representing the products of each group's population share and wealth share; and $R$ is the overlap (residual) component, which captures inequality due to the intersection of group wealth distributions. Lambert and Aronson (1993) provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap. A positive $R$ thus indicates significant intersections among wealth distributions of different groups[@lambert_inequality_1993].

::: {#tbl-tbl4}

{{< include ../output/rural_urban_summary.txt >}}

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary statistics of urban and rural income and wealth, 2017 & 2019
:::

The summary statistics highlight significant and persistent disparities between urban and rural China. Urban households consistently exhibit higher mean incomes and wealth compared to rural households. Urban mean wealth was 3.48 times greater than rural mean wealth in 2017 and widened slightly to 3.78 times in 2019. Additionally, the Gini coefficient for wealth is consistently higher in rural areas (0.6945 in 2017, 0.6629 in 2019) compared to urban areas (0.6573 in 2017, 0.6165 in 2019), though Gini coefficients for both groups declined modestly between the two years.

::: {#tbl-urbanrural-decomp}

{{< include ../output/urban_rural_gini_decomp.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Decomposition of wealth Gini coefficient by urban–rural division, 2017 & 2019 (%)
:::

The decomposition results indicate that about 55–57% (55.4% in 2017, 56.8% in 2019) of total wealth inequality is attributable to within-group disparities among urban and rural households. The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34% (32.6% in 2017, 33.7% in 2019). The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% (12.0% in 2017, 9.5% in 2019) of total inequality. The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.

::: {#tbl-region-summary}

{{< include ../output/region_summary.txt >}}

All figures are reported in RMB at constant 2017 prices.

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Summary Statistics by Region, 2017 and 2019
:::

The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019. Notably, the mean wealth of West-area households surpassed that of the Middle region in 2019, although its median wealth remained lower. Concurrently, the West's Gini coefficient rose to become the highest among all regions in 2019 (0.6947).

::: {#tbl-region-decomp}

{{< include ../output/region_gini_decomp.txt >}}

Source: Author's calculation based on the 2017 and 2019 CHFS data.

Decomposition of Wealth Gini Coefficient by Region, 2017 and 2019 (%)
:::

The group decomposition shows that, similar to the urban-rural decomposition, the between-group component contributes over 30% to overall wealth inequality in both 2017 and 2019 (36.9% and 33.8% respectively). The between-group contribution saw a slight decrease from 2017 to 2019, with a similar increase in the residual component (from 32.0% to 34.5%). This might signal some modest convergence in household wealth between regions.

## Conclusion
This paper analyzed wealth inequality in China using the 2017 and 2019 CHFS data, finding a modest decline in the overall Gini coefficient (0.701 to 0.667), consistent with literature suggesting a post-2010 peak [@kanbur_great_2021; @zhang_survey_2021]. Factor decomposition confirmed housing assets as the primary driver (contributing over 72%), aligning with previous studies [@knight_why_2022; @tan_earnings_2017], while financial equity's contribution grew (from approximately 11% to 14%). Partitioned analysis highlighted housing's stronger impact in urban areas (around 75% vs. approximately 52% rural) and especially the East region (over 80% vs. approximately 58% Non-East), with divergent trends emerging between 2017-2019 (increasing contribution in the East, decreasing elsewhere). Group decomposition further revealed that while within-group inequality dominates (accounting for approximately 55-60%), substantial inequality stems from persistent urban-rural (around 33%) and regional (over 30%) divides.

These dynamics likely reflect intertwined factors: rapid urbanization, uneven real estate appreciation and speculation impacting housing's role; financial market development driving the rise of financial assets, potentially favoring wealthier urban households; and deep-rooted structural issues like historical policies, the Hukou system, and differential access to opportunities maintaining the significant urban-rural and regional gaps, despite potential slow convergence hinted by recent trends and policies [@kanbur_great_2021]. Consequently, addressing China's high wealth inequality requires a nuanced approach beyond national measures, necessitating context-specific policies such as targeted housing market regulation in the East, support for land and business development in rural and Non-East areas, and broader efforts towards financial inclusion across the country.

# References